Disease progression of 213 patients hospitalized with Covid-19 in the Czech Republic in March - October 2020: An exploratory analysis

### 2021-09-03 manuscript/manuscript.Rmd
Fixed bungled math
Forgotten commit
Exporting figures as tiff

fig_progression
```

@@

fig_markers
  #plot_marker_data(c("d_dimer" = "D-dimer", "IL_6" = "IL 6", "CRP" = "CRP"), log_transform = TRUE)
#plot_marker_data(c("d_dimer", "IL_6", "CRP"), log_transform = FALSE)

@@

models<-list(m.ihd,m.hd,m.hf,m.copd,m.asthma,m.old,m.dia,m.rd,m.ld,m.s,m.ob,m.hc,m.hi,m.la)
models2<-list(m2.ihd,m2.hd,m2.hf,m2.copd,m2.asthma,m2.old,m2.dia,m2.rd,m2.ld,m2.s,m2.ob,m2.hc,m2.hi,m2.la)
modelNames<-c('IHD','HD','HF','COPD','Asthma','LungD','Dia','RenalD','LiverD','Smoking','Obesity','HighCr','HighInr','LowAlb')
gtmp<-c()
for (i in rev(seq(along=models))) {
  ii<-length(models)-i+1
  m<-models[[i]]
  m2<-models2[[i]]
  ci<-confint(m)[2,]
  ci2<-confint(m2)[2,]
  gtmp<-rbind(gtmp,data.frame(est=coef(m)[2],low=ci[1],high=ci[2],name=modelNames[i],type='outcome'))
  gtmp<-rbind(gtmp,data.frame(est=-coef(m2)[2],low=-ci2[1],high=-ci2[2],name=modelNames[i],type='hcq'))
}
ggplot(gtmp,aes(y=name,x=est,xmin=low,xmax=high,color=type))+
  geom_vline(xintercept=0,color='blue')+
  geom_point(position=position_dodge(width=0.5))+
  geom_errorbarh(position=position_dodge(width=0.5),height= 0.5)+
  scale_x_continuous('Log odds ratio')+
  scale_y_discrete('')+
  scale_color_manual('Model of:',values=c('black','red'),labels=c('Death','No HCQ'))+
  theme(legend.background = element_rect(fill="gray95"),legend.position=c(.83,.7))
```

@@

```{r multiverseresultsdrugs, fig.height=7, fig.width=8, fig.cap="Estimates of model coefficients for association between treatments and main outcomes. Each row represents a model - Categorical 7/28 = Bayesian categorical regression for state at day 7/28, Bayes Cox = Bayesian version of the Cox proportional hazards model with a binary outcome, Cox (single) = frequentist Cox model with a binary outcome, Cox (competing) = frequentist Cox model using competing risks (as in Fig 1a), HMM A = Bayesian hidden-Markov model as in Fig 1b with predictors for rate groups, HMM B = Bayesian hidden-markov model as in Fig 1b with predictors for individual rates, HMM C = Bayesian hidden-Markov model as in Fig 1c. For frequentist models, we show maximum likelihood estimate and 95\\% confidence interval, for Bayesian models we show posterior mean and 95\\% credible interval. The estimands are either log odds-ratio (Categorical, HMM) or log hazard ratio (Cox variants). In all cases coefficient < 0 means better patient outcome in the treatment group. Vertical lines indicate zero (blue) and substantial increase or decrease with odds or hazard ratio of 3:2 or 2:3 (green). Additionally the factors the model adjusted for are listed - Site = the study site, admitted = Admitted for Covid-19, Supportive = best supportive care initiated, Comorb. = total number of comorbidities, AZ = took azithromycin, HCQ = took hydroxychloroquine, FPV = took favipiravir, C. plasma = received convalescent plasma.  "}

@@

((multiverse_hcq1 & hide_axis_theme) / (multiverse_az1 & theme(axis.title.x = element_blank())) / multiverse_fpv1)  + plot_layout(heights = c(0.8, 0.7,0.70)) & theme(axis.title.y = element_blank())

@@

compare_auc_to_orig(all_auc_orig,
                    all_auc, show_outcome = FALSE)

@@

# Data availability

The data that support the findings of this study are available on request from the corresponding author or the secretariat of the Institute of Microbiology of the Czech Academy of Sciences (contact via mbu@biomed.cas.cz) for researchers who meet the criteria for access to confidential data. The data are not publicly available due to privacy restrictions imposed by the Ethical committee of General University Hospital in Prague and the GDPR regulation of the European Union.  We will be happy to arrange to run any analytical code locally and share the results, provided the code and the results do not leak personal information.

# Acknowledgements

This work was supported by ELIXIR CZ research infrastructure project (MEYS Grant No: LM2018131) including access to computing and storage facilities.

@@

* **S6 Fig. Association of HCQ with outcomes.** Estimates of model coefficients for association between hydroxychloroquine and main outcomes. The "Suspicious" section shows models that were found to not fit the data well or have computational issues - see supplementary statistical analysis for details.  Each row represents a model - Categorical All/7/28 = Bayesian categorical regression for state at last observed day/day 7/day 28, Binary All/7/28 = Bayesian logistic regression for state at last observed day/day 7/day 28, Bayes Cox = Bayesian version of the Cox proportional hazards model with a binary outcome, Cox (single) = frequentist Cox model with a binary outcome, Cox (competing) = frequentist Cox model using competing risks (as in Fig 1a), HMM A = Bayesian hidden-Markov model as in Fig 1b with predictors for rate groups, HMM B = Bayesian hidden-markov model as in Fig 1b with predictors for individual rates, HMM C = Bayesian hidden-Markov model as in Fig 1c. For frequentist models, we show maximum likelihood estimate and 95% confidence interval, for Bayesian models we show posterior mean and 95% credible interval. The estimands are either log odds-ratio (Categorical, HMM) or log hazard ratio (Cox variants) or log ratio of mean duration of hospitalization (HMM duration). In all cases coefficient < 0 means better patient outcome in the treatment group. Vertical lines indicate zero (blue) and substantial increase or decrease with odds or hazard ratio of 3:2 or 2:3 (green). Additionally the factors the model adjusted for are listed - Site = the study site, admitted = Admitted for Covid-19, Supportive = best supportive care initiated, Comorb. = total number of comorbidities, AZ = took azithromycin, HCQ = took hydroxychloroquine, FPV = took favipiravir, C. plasma = received convalescent plasma, first wave = only patients admitted before September 1st were included.

* **S7 Fig. Association of azithromycin with outcomes.** Estimates of model coefficients for association between azithromycin and main outcomes. The "Suspicious" section shows models that were found to not fit the data well or have computational issues - see supplementary statistical analysis for details.  Each row represents a model - Categorical All/7/28 = Bayesian categorical regression for state at last observed day/day 7/day 28, Binary All/7/28 = Bayesian logistic regression for state at last observed day/day 7/day 28, Bayes Cox = Bayesian version of the Cox proportional hazards model with a binary outcome, Cox (single) = frequentist Cox model with a binary outcome, Cox (competing) = frequentist Cox model using competing risks (as in Fig 1a), HMM A = Bayesian hidden-Markov model as in Fig 1b with predictors for rate groups, HMM B = Bayesian hidden-markov model as in Fig 1b with predictors for individual rates, HMM C = Bayesian hidden-Markov model as in Fig 1c. For frequentist models, we show maximum likelihood estimate and 95% confidence interval, for Bayesian models we show posterior mean and 95% credible interval. The estimands are either log odds-ratio (Categorical, HMM) or log hazard ratio (Cox variants) or log ratio of mean duration of hospitalization (HMM duration). In all cases coefficient < 0 means better patient outcome in the treatment group. Vertical lines indicate zero (blue) and substantial increase or decrease with odds or hazard ratio of 3:2 or 2:3 (green). Additionally the factors the model adjusted for are listed - Site = the study site, admitted = Admitted for Covid-19, Supportive = best supportive care initiated, Comorb. = total number of comorbidities, AZ = took azithromycin, HCQ = took hydroxychloroquine, FPV = took favipiravir, C. plasma = received convalescent plasma, first wave = only patients admitted before September 1st were included.

* **S8 Fig. Association of favipiravir with outcomes.** Estimates of model coefficients for association between favipiravir and main outcomes. The "Suspicious" section shows models that were found to not fit the data well or have computational issues - see supplementary statistical analysis for details.  Each row represents a model - Categorical All/7/28 = Bayesian categorical regression for state at last observed day/day 7/day 28, Binary All/7/28 = Bayesian logistic regression for state at last observed day/day 7/day 28, Bayes Cox = Bayesian version of the Cox proportional hazards model with a binary outcome, Cox (single) = frequentist Cox model with a binary outcome, Cox (competing) = frequentist Cox model using competing risks (as in Fig 1a), HMM A = Bayesian hidden-Markov model as in Fig 1b with predictors for rate groups, HMM B = Bayesian hidden-markov model as in Fig 1b with predictors for individual rates, HMM C = Bayesian hidden-Markov model as in Fig 1c. For frequentist models, we show maximum likelihood estimate and 95% confidence interval, for Bayesian models we show posterior mean and 95% credible interval. The estimands are either log odds-ratio (Categorical, HMM) or log hazard ratio (Cox variants) or log ratio of mean duration of hospitalization (HMM duration). In all cases coefficient < 0 means better patient outcome in the treatment group. Vertical lines indicate zero (blue) and substantial increase or decrease with odds or hazard ratio of 3:2 or 2:3 (green). Additionally the factors the model adjusted for are listed - Site = the study site, admitted = Admitted for Covid-19, Supportive = best supportive care initiated, Comorb. = total number of comorbidities, AZ = took azithromycin, HCQ = took hydroxychloroquine, FPV = took favipiravir, C. plasma = received convalescent plasma, first wave = only patients admitted before September 1st were included.

>>>>

fig_progression
ggsave(fig_progression, filename = "supp_figs_production/fig2.tiff", width = 7.5)
```

@@

fig_markers

ggsave(fig_markers, filename = "supp_figs_production/fig3.tiff", width = 7.5)
  #plot_marker_data(c("d_dimer" = "D-dimer", "IL_6" = "IL 6", "CRP" = "CRP"), log_transform = TRUE)
#plot_marker_data(c("d_dimer", "IL_6", "CRP"), log_transform = FALSE)

@@

models<-list(m.ihd,m.hd,m.hf,m.copd,m.asthma,m.old,m.dia,m.rd,m.ld,m.s,m.ob,m.hc,m.hi,m.la)
models2<-list(m2.ihd,m2.hd,m2.hf,m2.copd,m2.asthma,m2.old,m2.dia,m2.rd,m2.ld,m2.s,m2.ob,m2.hc,m2.hi,m2.la)
modelNames<-c('IHD','HD','HF','COPD','Asthma','LungD','Dia','RenalD','LiverD','Smoking','Obesity','HighCr','HighInr','LowAlb')
gtmp<-c()
for (i in rev(seq(along=models))) {
  ii<-length(models)-i+1
  m<-models[[i]]
  m2<-models2[[i]]
  ci<-confint(m)[2,]
  ci2<-confint(m2)[2,]
  gtmp<-rbind(gtmp,data.frame(est=coef(m)[2],low=ci[1],high=ci[2],name=modelNames[i],type='outcome'))
  gtmp<-rbind(gtmp,data.frame(est=-coef(m2)[2],low=-ci2[1],high=-ci2[2],name=modelNames[i],type='hcq'))
}
fig_assoc <- ggplot(gtmp,aes(y=name,x=est,xmin=low,xmax=high,color=type))+
  geom_vline(xintercept=0,color='blue')+
  geom_point(position=position_dodge(width=0.5))+
  geom_errorbarh(position=position_dodge(width=0.5),height= 0.5)+
  scale_x_continuous('Log odds ratio')+
  scale_y_discrete('')+
  scale_color_manual('Model of:',values=c('black','red'),labels=c('Death','No HCQ'))+
  theme(legend.background = element_rect(fill="gray95"),legend.position=c(.83,.7))

fig_assoc 
ggsave(fig_assoc, filename = "supp_figs_production/fig4.tiff", width = 7, height = 4.5)
```

@@

```{r multiverseresultsdrugs, fig.height=7, fig.width=8, fig.cap="Estimates of model coefficients for association between treatments and main outcomes. Each row represents a model - Categorical 7/28 = Bayesian categorical regression for state at day 7/28, Bayes Cox = Bayesian version of the Cox proportional hazards model with a binary outcome, Cox (single) = frequentist Cox model with a binary outcome, Cox (competing) = frequentist Cox model using competing risks (as in Fig 1a), HMM A = Bayesian hidden-Markov model as in Fig 1b with predictors for rate groups, HMM B = Bayesian hidden-markov model as in Fig 1b with predictors for individual rates, HMM C = Bayesian hidden-Markov model as in Fig 1c. For frequentist models, we show maximum likelihood estimate and 95\\% confidence interval, for Bayesian models we show posterior mean and 95\\% credible interval. The estimands are either log odds-ratio (Categorical, HMM) or log hazard ratio (Cox variants). In all cases coefficient $< 0$ means better patient outcome in the treatment group. Vertical lines indicate zero (blue) and substantial increase or decrease with odds or hazard ratio of 3:2 or 2:3 (green). Additionally the factors the model adjusted for are listed - Site = the study site, admitted = Admitted for Covid-19, Supportive = best supportive care initiated, Comorb. = total number of comorbidities, AZ = took azithromycin, HCQ = took hydroxychloroquine, FPV = took favipiravir, C. plasma = received convalescent plasma.  "}

@@

fig_drugs <- ((multiverse_hcq1 & hide_axis_theme) / (multiverse_az1 & theme(axis.title.x = element_blank())) / multiverse_fpv1)  + plot_layout(heights = c(0.8, 0.7,0.70)) & theme(axis.title.y = element_blank())

fig_drugs

ggsave(fig_drugs, filename = "supp_figs_production/fig5.tiff", width = 7.5, height = 7 * (7.5/8))
```

@@

fig_risk <- compare_auc_to_orig(all_auc_orig,
                    all_auc, show_outcome = FALSE)

fig_risk
ggsave(fig_risk, filename = "supp_figs_production/fig6.tiff", width = 7.5)
```

@@

@@

* **S6 Fig. Association of HCQ with outcomes.** Estimates of model coefficients for association between hydroxychloroquine and main outcomes. The "Suspicious" section shows models that were found to not fit the data well or have computational issues - see supplementary statistical analysis for details.  Each row represents a model - Categorical All/7/28 = Bayesian categorical regression for state at last observed day/day 7/day 28, Binary All/7/28 = Bayesian logistic regression for state at last observed day/day 7/day 28, Bayes Cox = Bayesian version of the Cox proportional hazards model with a binary outcome, Cox (single) = frequentist Cox model with a binary outcome, Cox (competing) = frequentist Cox model using competing risks (as in Fig 1a), HMM A = Bayesian hidden-Markov model as in Fig 1b with predictors for rate groups, HMM B = Bayesian hidden-markov model as in Fig 1b with predictors for individual rates, HMM C = Bayesian hidden-Markov model as in Fig 1c. For frequentist models, we show maximum likelihood estimate and 95% confidence interval, for Bayesian models we show posterior mean and 95% credible interval. The estimands are either log odds-ratio (Categorical, HMM) or log hazard ratio (Cox variants) or log ratio of mean duration of hospitalization (HMM duration). In all cases coefficient $< 0$ means better patient outcome in the treatment group. Vertical lines indicate zero (blue) and substantial increase or decrease with odds or hazard ratio of 3:2 or 2:3 (green). Additionally the factors the model adjusted for are listed - Site = the study site, admitted = Admitted for Covid-19, Supportive = best supportive care initiated, Comorb. = total number of comorbidities, AZ = took azithromycin, HCQ = took hydroxychloroquine, FPV = took favipiravir, C. plasma = received convalescent plasma, first wave = only patients admitted before September 1st were included.

* **S7 Fig. Association of azithromycin with outcomes.** Estimates of model coefficients for association between azithromycin and main outcomes. The "Suspicious" section shows models that were found to not fit the data well or have computational issues - see supplementary statistical analysis for details.  Each row represents a model - Categorical All/7/28 = Bayesian categorical regression for state at last observed day/day 7/day 28, Binary All/7/28 = Bayesian logistic regression for state at last observed day/day 7/day 28, Bayes Cox = Bayesian version of the Cox proportional hazards model with a binary outcome, Cox (single) = frequentist Cox model with a binary outcome, Cox (competing) = frequentist Cox model using competing risks (as in Fig 1a), HMM A = Bayesian hidden-Markov model as in Fig 1b with predictors for rate groups, HMM B = Bayesian hidden-markov model as in Fig 1b with predictors for individual rates, HMM C = Bayesian hidden-Markov model as in Fig 1c. For frequentist models, we show maximum likelihood estimate and 95% confidence interval, for Bayesian models we show posterior mean and 95% credible interval. The estimands are either log odds-ratio (Categorical, HMM) or log hazard ratio (Cox variants) or log ratio of mean duration of hospitalization (HMM duration). In all cases coefficient $< 0$ means better patient outcome in the treatment group. Vertical lines indicate zero (blue) and substantial increase or decrease with odds or hazard ratio of 3:2 or 2:3 (green). Additionally the factors the model adjusted for are listed - Site = the study site, admitted = Admitted for Covid-19, Supportive = best supportive care initiated, Comorb. = total number of comorbidities, AZ = took azithromycin, HCQ = took hydroxychloroquine, FPV = took favipiravir, C. plasma = received convalescent plasma, first wave = only patients admitted before September 1st were included.

* **S8 Fig. Association of favipiravir with outcomes.** Estimates of model coefficients for association between favipiravir and main outcomes. The "Suspicious" section shows models that were found to not fit the data well or have computational issues - see supplementary statistical analysis for details.  Each row represents a model - Categorical All/7/28 = Bayesian categorical regression for state at last observed day/day 7/day 28, Binary All/7/28 = Bayesian logistic regression for state at last observed day/day 7/day 28, Bayes Cox = Bayesian version of the Cox proportional hazards model with a binary outcome, Cox (single) = frequentist Cox model with a binary outcome, Cox (competing) = frequentist Cox model using competing risks (as in Fig 1a), HMM A = Bayesian hidden-Markov model as in Fig 1b with predictors for rate groups, HMM B = Bayesian hidden-markov model as in Fig 1b with predictors for individual rates, HMM C = Bayesian hidden-Markov model as in Fig 1c. For frequentist models, we show maximum likelihood estimate and 95% confidence interval, for Bayesian models we show posterior mean and 95% credible interval. The estimands are either log odds-ratio (Categorical, HMM) or log hazard ratio (Cox variants) or log ratio of mean duration of hospitalization (HMM duration). In all cases coefficient $< 0$ means better patient outcome in the treatment group. Vertical lines indicate zero (blue) and substantial increase or decrease with odds or hazard ratio of 3:2 or 2:3 (green). Additionally the factors the model adjusted for are listed - Site = the study site, admitted = Admitted for Covid-19, Supportive = best supportive care initiated, Comorb. = total number of comorbidities, AZ = took azithromycin, HCQ = took hydroxychloroquine, FPV = took favipiravir, C. plasma = received convalescent plasma, first wave = only patients admitted before September 1st were included.
